Algorithm explanation

### 2013-08-28 article/experiment.tex
Experiment

We have incorporated our algorithm to the software model checker for
higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It is
specifically designed to verify the correctness of higher-orde
programs written in a small subset of ML.  This verification tool
abstracts a given input program by inferring dependent types for
higher-order programs \cite{conf/ppdp/UnnoK09}.

We implemented our proposed algorithm into MoCHi by two steps.  First,
we replace an existing theorem prover with our interpolating theorem
prover as a sub-routine for solving Horn clauses in dependend typing.
We used a method to obtain a common small interpolant across different
interpolating problems, while reducing a Horn clause problem into
interpolation \cite{conf/ppdp/UnnoK09}.  Second, we completely
replaced MoCHi's Horn solving algorithm with our algorithm.

@@

While solving interpolating problems, our algorithm treated linear
equalities in the form $\mathbf{ax} + b = 0$ specially without
normalizing into a conjunction of linear equalities so that the
algorithm obtain a linear equality in a set of interpolants.  It is
because we observed some cases that requires a equality relation
between two variables but not an inequality as an program invariant.

@@

For linear constraint satisfaction, we used GNU Linear Programming Kit
(GLPK) first for its light computation load.  For Horn clause solving
however, Our implementation make use of Z3 \cite{conf/tacas/MouraB08}
as an SMT solver to obtain unsatisfiable core of the constraints.

@@

There are some interpolating theorem provers known.
\begin{itemize}
\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
  arithmetic reasonings and SAT-based logical reasonings.
\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
  proof-based method.
\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
 from constraints by linear arithmetic reasonings. It only supports
 conjunctive problems.
\end{itemize}

In our experiment, we compared the performance of our algorithm with
CSIsat, that is mainly used as a subprocedure of MoCHi.  Note that
CSIsat may return different interpolants for an interpolating problem
because of the randomized algorithm.

>>>>

We have incorporated our algorithm to the software model checker for
higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It is
specifically designed to verify the correctness of higher-order
programs written in a small subset of ML.  This verification tool
abstracts a given input program by inferring dependent types for
higher-order programs \cite{conf/ppdp/UnnoK09}.

We implemented our proposed algorithm into MoCHi by two steps.  First,
we replace an existing theorem prover with our interpolating theorem
prover as a sub-routine for solving Horn clauses in dependent typing.
We used a method to obtain a common small interpolant across different
interpolating problems, while reducing a Horn clause problem into
interpolation \cite{conf/ppdp/UnnoK09}.  Second, we completely
replaced MoCHi's Horn solving algorithm with our algorithm.

@@

While solving interpolating problems, our algorithm treated linear
equalities in the form $\mathbf{ax} + b = 0$ specially without
normalizing into a conjunction of linear equalities so that the
algorithm obtain a linear equality in a set of interpolants.  It is
because we observed some cases that require an equality relation
between two variables but not an inequality as an program invariant.

@@

For linear constraint satisfaction, we used GNU Linear Programming Kit
(GLPK) first for its light computation load.  For Horn clause solving
however, our implementation make use of Z3 \cite{conf/tacas/MouraB08}
as an SMT solver to obtain unsatisfiable core of the constraints.

@@

There are some interpolating theorem provers known.
\begin{itemize}
\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
  arithmetic reasoning and SAT-based logical reasoning.
\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
  proof-based method.
\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
 from constraints by linear arithmetic reasoning. It only supports
 conjunctive problems.
\end{itemize}

In our experiment, we compared the performance of our algorithm with
CSIsat, that is mainly used as a sub procedure of MoCHi.  Note that
CSIsat may return different interpolants for an interpolating problem
because of the randomized algorithm.

\paragraph{Comparison over simple interpolation}
First, we compare by simply generating an interpolant without
incorporating into MoCHi.  \textsc{Yint} represents our interpolating
theorem prover.

\begin{itemize}
\item Disjunctive interpolation in $A$.
\[\begin{array}{r l}
\varphi_A & (x-1 \geq 0) \vee (x+1 \leq 0) \\
\hline
\varphi_B & x=0
\end{array}\]
The result for this:
\[\begin{array}{l c}
\text{\textsc{Yint}} & (x+1 \leq 0) \vee (-x+1 \leq 0) \\
\text{\textsc{CSIsat}} & (-x \leq -1 \vee x \leq -1)
\end{array}\]
\item Disjunctive interpolation in $B$.
\[\begin{array}{r l}
\varphi_A & x=0 \\
\hline
\varphi_B & (x-1 \geq 0) \vee (x+1 \leq 0)
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & -x=0 \\
\text{\textsc{CSIsat}} & (-x \leq 0 \wedge x \leq 0)
\end{array}\]
\item Complex formula in $A$.
% complicated disj
\[\begin{array}{r l}
\varphi_A & ((x-5<0) \wedge (-x<0)) \vee ((x<0) \wedge (-x-5<0)) \\
\hline
\varphi_B & x=0
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & (x+1 \leq 0) \vee (-x+1 \leq 0) \\
\text{\textsc{CSIsat}} & (x < 0 \vee -x < 0)
\end{array}\]
\end{itemize}
For those problems that are simple enough, Yint and CSIsat returns the
same results.

Next we tried the problem that an equality interpolant exists.
% equality
\[\begin{array}{r l}
\varphi_A & x=0 \\
\hline
\varphi_B & x-1=0
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & x=0 \\
\text{\textsc{CSIsat}} & x \leq 0
\end{array}\]

\bigskip\centerline{}

% quantifier
\[\begin{array}{r l}
\varphi_A & (y-z=0) \wedge (x-z=0) \\
\hline
\varphi_B & (-w+y+1 \leq 0) \wedge (-w+x \geq 0)
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & x-y=0 \\
\text{\textsc{CSIsat}} & x-y \leq 0
\end{array}\]
Because our algorithm handles linear equalities special, our algorithm makes an effort to return an interpolant of linear equality if one exists.

Additionally, as the follwing resuls of \textsc{Yint} shows, if a simple interpolant exists for a given problem, our algorithm tries such result.
\[\begin{array}{r l}
\varphi_A & ((-y+1 \leq 0) \wedge (-x \leq 0)) \vee ((-y \leq 0) \wedge (-x+1 \leq 0)) \\
\hline
\varphi_B & ((y+1<0) \wedge (x<0)) \vee ((y<0) \wedge (x+1<0))
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & -x \leq 0 \\
\text{\textsc{CSIsat}} & (((x < 0 \vee -x \leq 0) \wedge 0 \leq x) \vee ((y < 0 \vee -y \leq 0) \wedge 0 \leq y))
\end{array}\]

\bigskip\centerline{}

\[\begin{array}{r l}
\varphi_A & (x-1=0) \vee (x+1=0) \\
\hline
\varphi_B & (x-2=0) \vee (x=0) \vee (x+2=0)
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & (-x-1=0) \vee (-x+1=0) \\
\text{\textsc{CSIsat}} & ((-x \leq -1 \vee x \leq -1) \wedge (-x \leq 1 \vee - x \leq - 1) \wedge (x \leq - 1 \vee x \leq 1))
\end{array}\]

Finally, we showed some test cases that may occur in practical interpolation.

\[\begin{array}{r l}
\multirow{2}{*}{$\varphi_A$}
& ((x-y-1 \leq 0) \wedge (-x+y-2 \leq 0) \wedge (x+y-1 \leq 0) \wedge (-x-y-2 \leq 0)) \vee \\
& ((x-y-2 \leq 0) \wedge (-x+y-1 \leq 0) \wedge (x+y-2 \leq 0) \wedge (-x-y-1 \leq 0)) \\
\hline
\multirow{2}{*}{$\varphi_B$}
& ((x-y+1<0) \wedge (-x+y-2 \leq 0) \wedge (x+y-2 \leq 0) \wedge (-x-y+1<0)) \vee \\
& ((x-y-2 \leq 0) \wedge (-x+y+1<0) \wedge (x+y+1<0) \wedge (-x-y-2 \leq 0))
\end{array}\]
\[\begin{array}{l c}
\text{\textsc{Yint}} & (-x+17y-33 \leq 0) \wedge (x-17y-33 \leq 0) \\
\text{\textsc{CSIsat}} & ((-1 \leq (x + y) \vee (x + -y) \leq 1) \wedge (-1 \leq (x + -y) \vee (x + y) \leq 1))
\end{array}\]

\bigskip\centerline{}

\[\begin{array}{r l}
\multirow{2}{*}{$\varphi_A$}
& ((y+1=0) \wedge (x-1=0)) \vee ((y+1=0) \wedge (x+1=0)) \vee \\
& ((y-1=0) \wedge (x+1=0)) \vee ((y-1=0) \wedge (x-1=0)) \\
\hline
\multirow{2}{*}{$\varphi_B$}
& ((y+2 \leq 0) \wedge (x-2 \geq 0)) \vee ((y+2 \leq 0) \wedge (x+2 \leq 0)) \vee \\
& ((y-2 \geq 0) \wedge (x+2 \leq 0)) \vee ((y-2 \geq 0) \wedge (x-2 \geq 0))
\end{array}\]
\begin{tabular}{l c}
\textsc{Yint} & $(-2y-3 \leq 0) \wedge (2y-3 \leq 0)$
\end{tabular}

\bigskip\centerline{}

\textsc{CSIsat}
\begin{align*}
& ((-y \leq 1 \wedge y \leq -1) \vee (-y \leq -1 \wedge x \leq 1) \vee x \leq -1) \wedge \\
& \qquad (-x \leq 1 \vee -x \leq -1 \vee -y \leq -1) \wedge \\
& \qquad (-x \leq 1 \vee -x \leq -1 \vee y \leq -1) \wedge \\
& \qquad (-x \leq -1 \vee x \leq -1)
\end{align*}

\[\begin{array}{r l}
\varphi_A & ((y-1=0) \wedge (x-1=0)) \vee ((y=0) \wedge (x=0)) \\
\hline
\varphi_B & (((y-1>0) \vee (y-1<0)) \wedge (x-1=0)) \vee (((y>0) \vee (y<0)) \wedge (x=0))
\end{array}\]
\begin{tabular}{l c}
\textsc{Yint} & $x-y=0$
\end{tabular}

\textsc{CSIsat}
\begin{align*}
& ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
& \qquad (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
& \qquad (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
& \qquad ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
& \qquad (1 = x \vee y \leq 0)
\end{align*}

### 2013-08-28 article/horn.tex
Experiment

The construction of the algorithm roughly consists of two parts.
\begin{itemize}
\item \textbf{Iterative Constraint Generation} is a process to build a
  constraint to compute a solution.  It starts from generating the
  weakest constraint in the sense that the resulting solution would
  not be simple yet.  The process iteratively strengthen a constraint
  until the strongest one while preserving the constraint
  satisfiability.  The strongest constraint gives a simple solution.
\item \textbf{Problem Restructuring} is a process to restructure the
  problem to obtain a less simpler solution, if the constraint become
  unsatisfiable during the iterative constraint generation.  This
  process considers the unsatisfiable core of the constraint, and
  wisely enlarge the solution size.
\end{itemize}

All predicate variables in an input problem are assigned simple
templates.  Those predicate templates are parameterized.  In the
iterative constraint generation process, algorithm builds a linear
constraint that parameters in templates must satisfy.  This is done in
a constructive manner, and constraints are build in the order of
implication dependency in Horn clauses.  That is, when an input
problem has a Horn clause thath has a predicate variable $P$ on its
left-hand and $Q$ on its right-hand, a constraint to describe
parameters in $P$'s template are first build before those in $Q$'s
template.  When all parameters in the whole problem are described, the
built constraint are evaluated to obtain a model.  Its instanciation
becomes an appropriate solution for the input problem.

The novelty of our algorithm lies in the incremental constraint
refinement.  The algorithm tries to solve from the weakest constraint
that an input problem allows that does not give a simple solution, and
gradually strengthen the constraint until it reaches to the simplest
solution that the problem has.  When a constraint becomes
unsatisfiable during this step, the algorithm analyzes the
unsatisfiable core and transform the problem to obtain a less simple
solution.

@@

Sinve we have $A$'s appearance twice on the second clause, it is
possible for $A$ to have two expressions each of which is used for
$A(x)$ and $A(y)$.  Although we need to use the same template for
multiple occurrences in order to compute the simplest solution, it may
make the whole constraints unsatisfiable if the simplest solution does
not exist.

@@

Because the unsatisfiable core contains constraints originating from
the disjunctions to be expressed in a single term template.  This let
us determine to relax the template to allow a enlarged solution.

@@

\begin{align*}
\text{Constraint:} \\
TODO \\
\\
\text{Unsatisfiable core:} \\
TODO
\end{align*}

@@

TODO

@@

Finally, we obtain the following constraint:

TODO

The process also generates a predicate template for every predicate
variable that the solution will become.  under certain solution
template.

@@

For ease of discussion, our algorithm handles an input set of Horn
clauses by a \emph{Horn graph} that express the equivalent problem.
The graph is then used throughout the algorithm to represent the
problem.  A Horn graph is a labelled directed acyclic graph
$G=(V,E,\varphi)$.
\begin{itemize}
\item $V$ is a union of two disjoint sets of vertices; Horn term
  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex
  $u \in V_\Rightarrow$ has eactly one outgoing edge to some
  $v \in V_T$, denoted as $succ(u)$.
\item $E$ is a set of labelled edges, which are expressed in the form
  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
  variables.
\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
  term vertices, where L is a set of elements of a predicate variable
  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
\end{itemize}
There are two kinds of edges in a Horn graph;
\begin{itemize}
\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
  empty.
\end{itemize}

@@

A variable mapping label $\theta$ on edges is used to map the index of
one predicate variable to another. (TODO?)

(Example?)

@@

\paragraph{Conjunction split}
For the first case previously mentioned, the unsatisfiable core
$\mathcal{U}$ contains coeffcient constraint expressions with a
coefficient parameter of certain predicate variable.  If the predicate
variable $P$ is the cause of unsatisfiability, some constraint
expressions in the unsatisfiable core have a parameter $a_{P,i}$
and/or $b_P$ in the following way.
\begin{align*}
% TODO:
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

Note that after this duplication of the predicate vertex $P$, in order
to induce different linear expressions for $P_1, \ldots, P_n$, it may
be necessary for preceeding predicate variable vertices in $G$ to be
duplicated in the same manner as $P$. It is because that different
derivation of $P$'s implication relation may be required to get
different consequences.

@@

After this duplication of the predicate vertex $P$, unlike the
conjunction split, this results in producing a disjunction at
succeeding predicate variables vertices ($R$ in this case).
\begin{align*}
(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
\end{align*}
Also, predicate variables that appeared with $P$ in the same
assumption ($Q$ in this case) needs to be duplicated by conjunction,
so as to obtain different premises that go along with $P_i$.
Therefore, the effect of disjunction split may reach to neighboring
predicate vertices.

@@

Conside example below. $e_i$ represents some conjunctions of linear
expressions.
% TODO: Can't we replace e_i with actual expressions?
\begin{align*}
(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
\end{align*}
In this example, a na\"{i}ve algorithm may solve following four
problems separately.
\begin{itemize}
\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
\end{itemize}
Then the solution is built by follows.
\begin{itemize}
\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
\end{itemize}
However, this method is not efficient.

@@

The algorithm first execute this duplication on all predicate
variables with multiple appearances.  It is the weakest constraint
that the algorithm makes for the certain $G$.  Then, the algorithm
gradually reduces the number of vertices that are duplicated by each
time testing the satisfiability of the constraint.  When the
constraint bebcome unsatisfiable at the certain iteration, the
algorithm can determine a predicate vertex which has to be split in a
conjunctive template.  If all the predicate variable constraints are
no longer duplicated and the constraint is the strongest, any model
for such constraint gives a simple solution under the Horn graph $G$
and the templateset $T$.

>>>>

The construction of the algorithm roughly consists of two parts.
\begin{itemize}
\item \textbf{Iterative Constraint Generation} is a process to build a
  constraint to compute a solution.  It starts from generating the
  weakest constraint in the sense that the resulting solution would
  not be simple yet.  The process iteratively strengthens a constraint
  until the strongest one while preserving the constraint
  satisfiability.  The strongest constraint gives a simple solution.
\item \textbf{Problem Restructuring} is a process to restructure the
  problem to obtain a less simple solution, if the constraint become
  unsatisfiable during the iterative constraint generation.  This
  process considers the unsatisfiable core of the constraint, and
  wisely enlarges the solution size.
\end{itemize}

All predicate variables in an input problem are assigned simple
templates.  Those predicate templates are parameterized.  In the
iterative constraint generation process, algorithm builds a linear
constraint that parameters in templates must satisfy.  This is done in
a constructive manner, and constraints are built in the order of
implication dependency in Horn clauses.  That is, when an input
problem has a Horn clause that has a predicate variable $P$ on its
left-hand and $Q$ on its right-hand, a constraint to describe
parameters in $P$'s template are first build before those in $Q$'s
template.  When all parameters in the whole problem are described, the
built constraint is evaluated to obtain a model.  Its instantiation
becomes an appropriate solution for the input problem.

The novelty of our algorithm lies in the incremental constraint
refinement.  The algorithm tries to solve from the weakest constraint
that an input problem allows that does not give a simple solution, and
gradually strengthen the constraint until it reaches to the simplest
solution that the problem has.  When a constraint becomes
unsatisfiable during this step, the algorithm analyzes the
unsatisfiable core and transforms the problem to obtain a less simple
solution.

@@

Since we have $A$'s appearance twice on the second clause, it is
possible for $A$ to have two expressions each of which is used for
$A(x)$ and $A(y)$.  Although we need to use the same template for
multiple occurrences in order to compute the simplest solution, it may
make the whole constraints unsatisfiable if the simplest solution does
not exist.

@@

Because the unsatisfiable core contains constraints originating from
the disjunctions to be expressed in a single term template, we can
determine which predicate to relax its template to allow an enlarged solution.

@@

\begin{align*}
\text{Constraint:} \\
& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
& c_1 + c_{3,2} = 0 \wedge c_2 + c_{4,2} > 0
\\
\text{Unsatisfiable core:} \\
0 \geq c_2 \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \\
& - \lambda_4 = c_{3,2} \wedge \\
& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
& c_1 + c_{3,2} = 0
\end{align*}

@@

@@

After this, we obtain the following constraint:

\begin{align*}
& \lambda_1 - \lambda_2 = c_{1,1} \wedge 0 \geq c_{2,1} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_1 - \lambda_2 = c_{1,2} \wedge 0 \geq c_{2,2} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
& c_{1,1} + c_{1,2} + c_{3,1} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,1} > 0 \wedge \\
& c_{1,1} + c_{1,2} + c_{3,2} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,2} > 0
\end{align*}

By continuing this process, the process in the end generates a
predicate template that the solution will become.  Eventually, we
obtain the following solution.

\begin{align*}
A(x) \text{ : } & x \leq 0 \wedge -x \leq 0 \\
B(x) \text{ : } & x+1 \leq 0 \vee -x+1 \leq 0
\end{align*}

@@

For ease of discussion, our algorithm handles an input set of Horn
clauses by a \emph{Horn graph} that express the equivalent problem.
The graph is then used throughout the algorithm to represent the
problem.  A Horn graph is a labeled directed acyclic graph
$G=(V,E,\varphi)$.
\begin{itemize}
\item $V$ is a union of two disjoint sets of vertices; Horn term
  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex
  $u \in V_\Rightarrow$ has exactly one outgoing edge to some
  $v \in V_T$, denoted as $succ(u)$.
\item $E$ is a set of labeled edges, which are expressed in the form
  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
  variables.
\item $\varphi: V_T \rightarrow L$ is a labeling function for Horn
  term vertices, where L is a set of elements of a predicate variable
  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
\end{itemize}
There are two kinds of edges in a Horn graph;
\begin{itemize}
\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
  empty.
\end{itemize}

@@

A variable mapping label $\theta$ on edges is used to map the index of
one predicate variable to another.

@@

\paragraph{Conjunction split}
For the first case previously mentioned, the unsatisfiable core
$\mathcal{U}$ contains coefficient constraint expressions with a
coefficient parameter of certain predicate variable.  If the predicate
variable $P$ is the cause of unsatisfiability, some constraint
expressions in the unsatisfiable core have a parameter $a_{P,i}$
and/or $b_P$ in the following way.
\begin{align*}
% TODO:
& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
& \left( \exists i, j, k;
\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
\right\rbrace \subseteq \mathcal{U} \right)
\end{align*}

@@

Note that after this duplication of the predicate vertex $P$, in order
to induce different linear expressions for $P_1, \ldots, P_n$, it may
be necessary for preceding predicate variable vertices in $G$ to be
duplicated in the same manner as $P$. It is because that different
derivation of $P$'s implication relation may be required to get
different consequences.

@@

After this duplication of the predicate vertex $P$, unlike the
conjunction split, a disjunction at succeeding predicate
variable vertices ($R$ in this case) is produced.
\begin{align*}
(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
\end{align*}
Also, predicate variables that appeared with $P$ in the same
assumption ($Q$ in this case) needs to be duplicated by conjunction,
so as to obtain different premises that go along with $P_i$.
Therefore, the effect of disjunction split may reach to neighboring
predicate vertices.

@@

Consider example below. $e_i$ represents some conjunctions of linear
expressions.
% TODO: Can't we replace e_i with actual expressions?
\begin{align*}
(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
\end{align*}
In this example, a na\"{i}ve algorithm may solve following four
problems separately.
\begin{itemize}
\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
\end{itemize}
Then the solution is built by follows.
\begin{itemize}
\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
\end{itemize}
However, this method is not efficient.

@@

The algorithm first execute this duplication on all predicate
variables with multiple appearances.  It is the weakest constraint
that the algorithm makes for the certain $G$.  Then, the algorithm
gradually reduces the number of vertices that are duplicated by each
time testing the satisfiability of the constraint.  When the
constraint becomes unsatisfiable at the certain iteration, the
algorithm can determine a predicate vertex which has to be split in a
conjunctive template.  If all the predicate variable constraints are
no longer duplicated and the constraint is the strongest, any model
for such constraint gives a simple solution under the Horn graph $G$
and the templateset $T$.
